### 2021-04-03 article/paper_eng.tex
changed email
fixed align
fixes

\begin{center}
    \huge\textbf{Training inhibitory connections between local receptive fields in spiking neural networks}\\
\end{center}

@@

\textsuperscript{(a)} Lomonosov Moscow State University, Moscow, Russia \linebreak
\textsuperscript{(b)} National Research Center ``Kurchatov Institute'', Moscow, Russia \linebreak
\textsuperscript{(\textasteriskcentered)} Corresponding author. Email address: \textit{danielgafni16@gmail.com}

@@

\section*{Acknowledgments}
This work was partially supported by the Russian Science Foundation (Project № 20-79-10185) in the studying of inhibitory connections training part and by the Russian Foundation for Basic Research (Project № 18-29-23041) in the part of comparison of convolutional and locally connected achitectures.

\addcontentsline{toc}{section}{Abstract}
\section*{Abstract}
Standard learning methods (based on the error backpropagation) used in artificial neural networks (ANNs) are difficult to apply to spiking neural networks (SNNs) due to discrete and time-distributed nature of the events they generate. Although such projects exist, local learning algorithms, which are more biologically correct, capable of performing unsupervised learning and more energy efficient (if implemented on specialized hardware), are of greater interest.

@@

The introduction of inhibitory connections (with negative weights) allows to achieve a better distinction of features learned by the neurons. In this work we show that learning competition relationships in a locally connected (LC) SNN potentially leads to higher pattern recognition accuracy due to the organization of competition and cooperation between groups of neurons. It is also shown that the LC SNN produces better results than a convolutional SNN.

@@

Modern abstract neural networks (represented by combinations of linear layer transformations and non-linear activation functions) do an excellent job at solving many machine learning problems \parencite{pmlr-v28-wan13, Khan_2020}. However, training them is a laborious process that requires large computational resources. Typically, training is conducted using hundreds or thousands of examples and can take months. Both training and inference of formal neural network algorithms are far from being effective \parencite{Edwards2015GrowingPF}. This is due both to the physically separate storage of connection weights and neuron activations, and to the calculations themselves, which are tensor in nature, associated with vector-matrix and matrix-matrix multiplication of floating point numbers. Modern CPUs are not optimized for this kind of computing. GPUs --- architectures originally created for working with computer graphics, and therefore more suitable for tensor calculations --- are much better at these tasks than traditional CPUs, but they do not perform with the desired efficiency either. The number of parameters in modern formal neural networks can reach tens \parencite{ManyParams, Khan_2020} and even hundreds of billions (the latest OpenAI language model GPT-3 \parencite{brown2020language}).

@@

\item Moreover, these algorithms can use unsupervised learning, which does not require manual data labeling, as in the case of backpropagation methods.

@@

\item SNNs, to a greater extent than formal neural networks, biologically correctly model the interactions of neurons, which can be used for biological simulations of the nervous system for the purposes of studying not only bioinformatics, but also biophysical and medical aspects of its functioning.
\end{itemize}

@@

\begin{figure}[H] \label{LCSNN}
\begin{center}
\includegraphics[,
 width=0.6\textwidth,keepaspectratio=true]{LCSNN.pdf} 
    \caption{LCSNN architecture}
    \label{fig:LC_SNN_architecture}
\end{center}
\end{figure}

@@

\subsection{Forward connections training}
To reduce the number of model parameters, images are cropped to 20 $\times$ 20 pixels. The edges of the images are often almost empty, so this operation has little or no effect on the amount of information, available to the network. For each image, $X$ spikes are generated using a Poisson distribution with a mathematical expectation value  proportional to the corresponding pixel intensiveness. Connections $XY$ are trained according to the STDP \parencite{STDP} rule. This is a biologically inspired rule of unsupervised learning. When the presynaptic spike (pre-spike) is received and the post-spike is emitted, the weight $w$ of the corresponding connection is increased by $ \Delta w $, where
\begin{equation} 
\Delta w =
 \begin{cases}
 A_+ \cdot e^{- \frac{t_{pre} - t_{post}}{\tau_+}}, t_{pre} - t_{post} > 0;\\
 A_- \cdot e^{- \frac{t_{pre} - t_{post}}{\tau_-}}, t_{pre} - t_{post} < 0.
 \end{cases}
\end{equation}

@@

Thus, in the process of learning, the weights of connections increase if pre-spikes are systematically recieved right before the emmision of post-spikes, and decrease in the opposite case. After such training, the neurons begins to actively react to pre-spikes received from neurons with larger connection weights. This leads to the neuron weights learning certain features (Fig. \refbf{fig:training}). If these pre-spikes are recieved in a short enough time period, they might cause a new spike. The opposite rule (with $A_{+}$ and $A_{-}$ having the opposite signs) is called the Anti-STDP \parencite{Anti-STDP} rule. It is used to train inhibitory connenctions in section \refbf{section:competition-training} of this paper.

@@

\begin{figure}[H]
\centering
\begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio=true]{weights_XY_untrained.pdf} 
    \caption{}
    \label{fig:training:before_training}
\end{subfigure}
\begin{subfigure}{0.45\textwidth}  \label{weights_XY}
    \includegraphics[width=\textwidth,keepaspectratio=true]{weights_XY.pdf}
    \caption{} 
    \label{fig:training:after_training}
\end{subfigure}
\caption{Visualization of $XY$ connection weights of a 900 $Y$ neurons (100 neurons for each of the 9 patches) network --- a 360 $ \times $ 360 matrix. The weights of еру one neuron correspond to a 12 $\times$ 12 square. The weights of the $Y$ neurons are grouped by the patch --- the lower left 120 $  \times $ 120 square  (10 $ \times $ 10 neurons) corresponds to the lower left patch, the center square corresponds to the center patch, and so on. Weights are initialized randomly (\subrefbf{fig:training:before_training}), and after 1000 training iterations (\subrefbf{fig:training:after_training}) the neurons learnin different features, resembling digits elements}
\label{fig:training}
\end{figure}

@@

\begin{figure}
\begin{subfigure}{0.49\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio=true]{LCSNN_learning_rate.pdf}
 \caption{}
 \label{LCSNN_learning_curve}
\end{subfigure}
\begin{subfigure}{0.49\textwidth} 
    \includegraphics[width=\textwidth,keepaspectratio=true]{CSNN_learning_rate.pdf}
 \caption{}
 \label{CSNN_learning_curve}
\end{subfigure} 
\begin{subfigure}{0.49\textwidth} 
    \includegraphics[width=\textwidth,keepaspectratio=true]{FCSNN_learning_rate.pdf}
 \caption{}
 \label{FCSNN_learning_curve}
\end{subfigure}
\caption{Comparison of learning curves of different architectures of SNNs with competition: (\subrefbf{LCSNN_learning_curve}) LCSNN, (\subrefbf{CSNN_learning_curve}) CSNN, (\subrefbf{FCSNN_learning_curve}) FCSNN. The results are evaluated with different algorithms for the interpretation of activity. The \textbf{preselections by spikes} algorithm was only checked for LCSNN, because it achieves practically same results as \textbf{patch voting}. Standard deviations are plotted as errors. The networks have 100 channels (for a FCSNN the channel consists of a single neuron), the convolution kernel size ($k$) for the LCSNN is 12, and 8 for the CSNN}
\label{fig:learning_curves}
\end{figure}

@@

It should be noted that SNNs can achieve significantly higher recognition accuracy on MNIST. It can be achieved by: (i) networks with a significantly larger number of parameters (for example, with an increased number of channels); (ii) deeper networks; (iii) more efficient learning algorithms (for example, supervised learning or using the contrastive loss function \parencite{contrastive_loss}). At the same time, achieving the highest possible accuracy was not the goal of this study, which is primarily aimed at comparing various SNN architectures.

@@

The results obtained in this work are comparable to the results in \parencite{saunders2019locally}. Note that the results achieved with unsupervised training are only slightly inferior to the results achived by supervised training. Note, that our best locally connected network with trainable competitions had $ 10^7 $ parameters (of which $ 1 \cdot 10^6 $ are forward connections and $ 9 \cdot 10^6 $ are inhibitory connections), which is less than $ 4.6 \cdot 10^7 $ (of which $ 0.5 \cdot 10^7 $ are forward conenctions and $ 4.1 \cdot 10^7 $ are compettition connections) for a fully connected network with competition from \parencite{mnist2}. In addition, the network from \parencite{mnist2} was trained for $ 1 \cdot 10^6 $ iterations, while in this study the number of training iterations did not exceed 5000.

\section{Inhibitory connections training}
\label{section:competition-training}
Inhibitory connections $YY$ greatly affect the training of forward connections $XY$. Large absolute values of inhibitory weights contibute to greater variability and specialization in training $Y$ neurons, as for each receptive field similar weights $XY$ neoruns are simultaneously active (Fig. \refbf{fig:competition_effect_on_training:high_comp}). On the contrary, inhibitory weights of small absolute values do not allow neurons to specialize effectively (Fig. \refbf{fig:competition_effect_on_training:low_comp}).

@@

All SNNs discussed in this paper up to this point had fixed competition. In this regard, the question arised of how the trianing of inhibitory weights affects the accuracy of the network.

@@

\begin{table}[H]
 \caption{Anti-STDP parameters}
\begin{center}
\begin{tabular}{|l|l|l|l|l|}
\hline
Figure & {$\tau_{+}$, ms} & {$\tau_{-}$, ms} & {$A_{+}$} & {$A_{-}$} \\
\hline
\refbf{fig:competition_distributions:weakest}, \refbf{fig:worst_competition_distribution} & 14.7 & 14.2 & -0.5 & 1.5 \\
\hline
\refbf{fig:competition_distributions:weak}, \refbf{fig:medium_bad_competition_distribution} & 5.4 & 15.1 & -1.2 & 0.6 \\
\hline
\refbf{fig:competition_distributions:medium}, \refbf{fig:medium_good_competition_distribution} & 17.6 & 24.5 & -1.9 & 1.6 \\
\hline
\refbf{fig:competition_distributions:strong}, \refbf{fig:best_competition_XY} & 17.7 & 16.5 & -0.1 & 1.6\\
\hline
\end{tabular}
\end{center}
\end{table}

\begin{figure}[H]
\centering
\begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio=true]{competition_distribution_worst.pdf}
    \caption{}
    \label{fig:competition_distributions:weakest}
\end{subfigure}
\begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio=true]{competition_distribution_medium_bad.pdf}
    \caption{}
    \label{fig:competition_distributions:weak}
\end{subfigure}
\\
\begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio=true]{competition_distribution_medium_good.pdf}
    \caption{}
    \label{fig:competition_distributions:medium}
\end{subfigure}
\begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio=true]{competition_distribution_best.pdf}
    \caption{}
    \label{fig:competition_distributions:strong}
\end{subfigure}
\caption{Different inhibitory ($YY$) weights distributions: weakest competition (\subrefbf{fig:competition_distributions:weakest}), weak competition (\subrefbf{fig:competition_distributions:weak}), medium competition (\subrefbf{fig:competition_distributions:medium}) and strong competition (\subrefbf{fig:competition_distributions:strong})}
\label{fig:competition_distributions}
\end{figure}

@@

During the optimization of the models, the accuracy of a large number of LCSNNs was measured with various hyperparameter configurations (Fig. \refbf{fig:hyperparams}). It can be seen that configurations with high accuracy (dark lines) are localized in a manner that encourages negative weight accumulation for larger time differencies $t_{post} - t_{pre}$ between -post and -pre spikes (for the Anti-STDP weight update rule).

@@

\subsection{Trained inhibitory weights analysis}

In addition, inhibitory weights clamping (during training) experiments were carried out. It turned out that the inhibitory weights in the entire range of their variation are important for the effective LCSNN functioning, since both the upper and the lower clamps negatively affect the classification accuracy (Fig. \refbf{fig:compe_clamp}). This is due to the fact that high competition contributes to greater specialization of neurons and therefore is useful, and low competition allows neurons to cooperate and recognize digits together (including ensuring the accumulation of more statistics for calibrating the voices of neurons).

\begin{figure}[H] 
\centering 
\begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio=true]{competition_distribution_clamp_low.pdf}
    \caption{}
    \label{fig:compe_clamp:low_distr}
\end{subfigure}
\begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio=true]{competition_distribution_clamp_high.pdf}
    \caption{}
    \label{fig:compe_clamp:high_distr}
\end{subfigure}
\begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio=true]{weights_XY_clamp_low.pdf}
    \caption{}
    \label{fig:compe_clamp:low_weights}
\end{subfigure} 
\begin{subfigure}{0.3\textwidth} 
    \includegraphics[width=\textwidth,keepaspectratio=true]{weights_XY_clamp_high.pdf}
    \caption{}
    \label{fig:compe_clamp:high_weights}
\end{subfigure}
\caption{Impact of limiting the values of inhibitory weights on accuracy. \subrefbf{fig:compe_clamp:low_weights}, \subrefbf{fig:compe_clamp:high_weights} show $XY$ weights and Figures (\subrefbf{fig:compe_clamp:low_distr}, \subrefbf{fig:compe_clamp:high_distr}) show the inhibitory ($YY$) weights distributions. The inhibitory weights were clamped (during training) to $[-350, 0]$ (\subrefbf{fig:compe_clamp:low_distr},\subrefbf{fig:compe_clamp:low_weights}) and to $(-\infty, -100]$ (\subrefbf{fig:compe_clamp:high_distr},\subrefbf{fig:compe_clamp:high_weights}). The accuracy of the network with the same hyperparameters, but without the competition limitation is 82.3\%, which is significantly higher than the accuracies of the networks presented in this figure. Its weights $ XY $ are shown in Fig. \refbf{fig:best_competition_XY}, and the distribution of its inhibitory weights is shown in Fig. \refbf{fig:competition_distributions:strong}. In both cases the networks develop more blurred features comparing to the network with unrestricted competition in Fig. \refbf{fig:competition_distributions:strong}
} 
\label{fig:compe_clamp}
\end{figure}

Competition training made it possible to achieve an accuracy slightly (by 2 \%) higher than the accuracy of a network of the same configuration, but with a fixed competition with inhibitory weights equal to $ -50 $ (Table \refbf{results}, №4 and №5).

\section{Discussion}
This study demonstrates that a locally connected network is a promising spiking neural network architecture suitable for efficient implementation on a specialized neural hardware. Indeed, the ability to quickly plateau the learning curve allows the LCSNN to be used for training on relatively small training samples (3000-5000). The use of an additional activity interpretation algorithm, such as a linear classifier, can improve the efficiency of the network. Note that in this case the very features of the objects being recognized are learned by the network in an unsupervised fashion. The question remains about the possibility of obtaining similar results without using supervised learning to interpret network activity.

The main advantage of LCSNN is, as has been shown, that the locally connected network outperforms the convolutional in accuracy with approximately the same or slightly larger number of parameters. Presumably, this is due to the richer statistics of patches (receptive fields) that neurons within each channel respond to, compared to one receptive field per channel in the convolutional architecture. The price for this is the increased number of network parameters, but is not a critical factor due to the local (as opposed to fully connected) topology of interneural contacts. The local architecture also allows to save information about the localization of the selected feature in space and therefore is a reasonable choice for the implementation of unsupervised trained SNNs. This choice also determines its overall efficiency (in terms of performance, accuracy and power consumption at a certain density of elements) in a hardware implementation, in which each weight is usually represented by one or more physical elements (mainly SRAM cells \parencite{TrueNorth, Loihi} or memristors \parencite{Li_2018}).
The training of inhibitory connections slightly increases the accuracy of the LCSNN-based algorithm, but it is not so critical as to justify the use of such an expensive operation as the introduction and training of a significant number of inhibitory connections with finite weights of different values. At the same time, this conclusion is probably only valid only for small networks, like the one studied in this paper. In our opinion, the selection of optimal algorithms and parameters for training competition for deep SNNs deserves a more thorough study, in which the balance of competition and cooperation of neurons within each layer can lead to the formation of clusters from classes of supplied images, which contributes to building a semantically-like hierarchy of learned features and their combinations, that is, the final images \parencite{10.1007/978-3-030-30425-6_30}. Potentially, such a hierarchy of images can lead to a significant increase in the quality metrics.

>>>>

\begin{center}
    \huge\textbf{Training Inhibitory Connections Between Local Receptive Fields in Spiking Neural Networks}\\
\end{center}

@@

\textsuperscript{(a)} Lomonosov Moscow State University, Moscow, Russia \linebreak
\textsuperscript{(b)} National Research Center ``Kurchatov Institute'', Moscow, Russia \linebreak
\textsuperscript{(\textasteriskcentered)} Corresponding author. Email address: \textit{daniilgafni@gmail.com}

@@

\section*{Acknowledgments}
This work was supported by the Russian Foundation for Basic Research (Project № 18-29-23041) in the comparison of convolutional and locally connected achitectures and by the Russian Science Foundation (Project № 20-79-10185) in the studying of inhibitory connection training part.

\addcontentsline{toc}{section}{Abstract}
\section*{Abstract}
Standard learning methods (based on the error backpropagation technique) used in artificial neural networks (ANNs) are difficult to apply to spiking neural networks (SNNs) due to asynchronous and continuous-time nature of the events they generate. Although such projects exist, local learning algorithms, which are more biologically plausible, capable of performing unsupervised learning and more energy efficient (if implemented on specialized hardware), are of greater interest.

@@

The introduction of inhibitory connections (with negative weights) allows to achieve a better distinction of features learned by the neurons. In this work we show that learning competition relationships in a locally connected (LC) SNN potentially leads to higher pattern recognition accuracy due to the organization of competition and simultaneously cooperation between groups of neurons. It is also shown that the LC SNN produces better results than a convolutional SNN, even at the same or lesser number of trainable synaptic weight parameters. This study proceeds paving the way towards building effective computational algorithms based on SNNs with biologically plausible architectures and life-long learning.

@@

Modern formal neural networks (represented by combinations of linear layer transformations and non-linear activation functions) do an excellent job at solving many machine learning problems \parencite{pmlr-v28-wan13, Khan_2020}. However, training them is a laborious process that requires large computational resources. Typically, training is conducted using hundreds or thousands of examples and can take months. Both training and inference of formal neural network algorithms are far from being effective \parencite{Edwards2015GrowingPF}. This is due both to the physically separate storage of connection weights and neuron activations, and to the calculations themselves, which are tensor in nature, associated with vector-matrix and matrix-matrix multiplication of floating point numbers. Modern CPUs are not optimized for this kind of computing. GPUs --- architectures originally created for working with computer graphics, and therefore more suitable for tensor calculations --- are much better at these tasks than traditional CPUs, but they do not perform with the desired efficiency either. The number of parameters in modern formal neural networks can reach tens \parencite{ManyParams, Khan_2020} and even hundreds of billions (the latest OpenAI language model GPT-3 \parencite{brown2020language}).

@@

\item Moreover, these algorithms can use unsupervised learning \parencite{nonlinearmemristivedynamics, stdpbasedclassification}, which does not require manual data labeling, as in the case of backpropagation methods.

@@

\item SNNs, to a greater extent than formal neural networks, biologically correctly model the interactions of neurons \parencite{hippocampus}, which can be used for biological simulations of the nervous system for the purposes of studying not only bioinformatics, but also biophysical and medical aspects of its functioning.
\end{itemize}

@@

\begin{figure}[H] \label{LCSNN}
\begin{center}
\includegraphics[
 width=0.6\textwidth,keepaspectratio=true]{LCSNN.pdf} 
    \caption{LCSNN architecture}
    \label{fig:LC_SNN_architecture}
\end{center}
\end{figure}

@@

\subsection{Forward connections training}
To reduce the number of model parameters, images are cropped to 20 $\times$ 20 pixels. The edges of the images are often almost empty, so this operation has little or no effect on the amount of information, available to the network. For each image, $X$ spikes are generated using a Poisson distribution with a mathematical expectation value  proportional to the corresponding pixel intensitiy. Connections $XY$ are trained according to the STDP \parencite{STDP} rule. This is a biologically inspired rule of unsupervised learning. When the presynaptic spike (pre-spike) is received and the post-spike is emitted, the weight $w$ of the corresponding connection is increased by $ \Delta w $, where
\begin{equation} 
\Delta w =
 \begin{cases}
 A_+ \cdot e^{- \frac{t_{post} - t_{pre}}{\tau_+}}, t_{post} - t_{pre} > 0;\\
 A_- \cdot e^{\frac{t_{post} - t_{pre}}{\tau_-}}, t_{post} - t_{pre} < 0.
 \end{cases}
\end{equation}

@@

Thus, in the process of learning, the weights of connections increase if pre-spikes are systematically recieved right before the emmision of post-spikes, and decrease in the opposite case. All pre- and post- spikes interact with each other in the connection weight update. After such training, the neurons begins to actively react to pre-spikes received from neurons with larger connection weights. This leads to the neuron weights learning certain features (Fig. \refbf{fig:training}). If these pre-spikes are recieved in a short enough time period, they might cause a new spike. The opposite rule (with $A_{+}$ and $A_{-}$ having the opposite signs) is called the Anti-STDP \parencite{Anti-STDP} rule. It is used to train inhibitory connenctions in section \refbf{section:competition-training} of this paper.

@@

\begin{figure}[H]
\centering
\begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio=true]{weights_XY_untrained.pdf} 
    \caption{}
    \label{fig:training:before_training}
\end{subfigure}
\begin{subfigure}{0.45\textwidth}  \label{weights_XY}
    \includegraphics[width=\textwidth,keepaspectratio=true]{weights_XY.pdf}
    \caption{} 
    \label{fig:training:after_training}
\end{subfigure}
\caption{Visualization of $XY$ connection weights of a 900 $Y$ neurons (100 neurons for each of the 9 patches) network --- a 360 $ \times $ 360 matrix. The weights of a one neuron correspond to a 12 $\times$ 12 square. The weights of the $Y$ neurons are grouped by the patch --- the lower left 120 $  \times $ 120 square  (10 $ \times $ 10 neurons) corresponds to the lower left patch, the center square corresponds to the center patch, and so on. Weights are initialized randomly (\subrefbf{fig:training:before_training}), and after 1000 training iterations (\subrefbf{fig:training:after_training}) the neurons learn different features resembling digit elements}
\label{fig:training}
\end{figure}

@@

\begin{figure}
\begin{subfigure}{0.49\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio=true]{LCSNN_learning_rate.pdf}
 \caption{}
 \label{LCSNN_learning_curve}
\end{subfigure}
\begin{subfigure}{0.49\textwidth} 
    \includegraphics[width=\textwidth,keepaspectratio=true]{CSNN_learning_rate.pdf}
 \caption{}
 \label{CSNN_learning_curve}
\end{subfigure} 
\begin{subfigure}{0.49\textwidth} 
    \includegraphics[width=\textwidth,keepaspectratio=true]{FCSNN_learning_rate.pdf}
 \caption{}
 \label{FCSNN_learning_curve}
\end{subfigure}
\caption{Comparison of different competition SNN architectures learning curves: (\subrefbf{LCSNN_learning_curve}) LCSNN, (\subrefbf{CSNN_learning_curve}) CSNN, (\subrefbf{FCSNN_learning_curve}) FCSNN. The results are evaluated with different algorithms for the interpretation of activity. The \textbf{preselections by spikes} algorithm was only checked for LCSNN, because it achieves practically the same results as \textbf{patch voting}. Standard deviations are plotted as errors. The networks have 100 channels (for a FCSNN the channel consists of a single neuron), the convolution kernel size ($k$) for the LCSNN is 12, and 8 for the CSNN}
\label{fig:learning_curves}
\end{figure}

@@

It should be noted that SNNs can achieve significantly higher recognition accuracy on MNIST. It can be achieved by: (i) networks with a significantly larger number of parameters (for example, with an increased number of channels); (ii) deeper networks; (iii) more efficient learning algorithms (for example, supervised learning or using the contrastive loss function \parencite{contrastive_loss}). At the same time, obtaining the highest possible accuracy was not the goal of this study, which is primarily aimed at comparing various SNN architectures.

@@

The results obtained in this work are comparable to the results in \parencite{saunders2019locally}, which are demonstrated by using additionally a kind of metabolic rule for weight sum normalization for each neuron. Note that the results achieved with unsupervised training are only slightly inferior to the results achived by supervised training. It is also worth noting, that our best locally connected network with trainable competitions had $ 10^7 $ parameters (of which $ 1 \cdot 10^6 $ are forward connections and $ 9 \cdot 10^6 $ are inhibitory connections), which is less than $ 4.6 \cdot 10^7 $ (of which $ 0.5 \cdot 10^7 $ are forward conenctions and $ 4.1 \cdot 10^7 $ are compettition connections) for a fully connected network with competition from \parencite{mnist2}. In addition, the network from \parencite{mnist2} was trained for $ 1 \cdot 10^6 $ iterations, while in this study the number of training iterations did not exceed 5000.

\section{Inhibitory connection training}
\label{section:competition-training}
Inhibitory connections $YY$ greatly affect the training of forward connections $XY$. Large absolute values of inhibitory weights contribute to greater variability and specialization in training $Y$ neurons, as for each receptive field similar weights $XY$ neoruns are simultaneously active (Fig. \refbf{fig:competition_effect_on_training:high_comp}). On the contrary, inhibitory weights of small absolute values do not allow neurons to specialize effectively (Fig. \refbf{fig:competition_effect_on_training:low_comp}).

@@

All SNNs discussed in this paper up to this point had a fixed competition. In this regard, the question arised of how the trianing of inhibitory weights affects the accuracy of the network.

@@

\begin{table}[H]
 \caption{Anti-STDP parameters. $\tau_{-}$ and $\tau_{+}$ affect both the Anti-STDP training of the $YY$ connections and the STDP trainig of $XY$ connections}
\begin{center}
\begin{tabular}{|l|l|l|l|l|}
\hline
Figure & {$\tau_{+}$, ms} & {$\tau_{-}$, ms} & {$A_{+}$} & {$A_{-}$} \\
\hline
\refbf{fig:competition_distributions:weakest}, \refbf{fig:worst_competition_distribution} & 14.7 & 14.2 & -1.5 & 0.5 \\
\hline
\refbf{fig:competition_distributions:weak}, \refbf{fig:medium_bad_competition_distribution} & 5.4 & 15.1 & -0.6 & 1.2 \\
\hline
\refbf{fig:competition_distributions:medium}, \refbf{fig:medium_good_competition_distribution} & 17.6 & 24.5 & -1.6 & 1.9 \\
\hline
\refbf{fig:competition_distributions:strong}, \refbf{fig:best_competition_XY} & 17.2 & 16.4 & -1.6 & 0.1\\
\hline
\end{tabular}
\end{center}
\end{table}

\begin{figure}[H]
\centering
\begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio=true]{competition_distribution_worst.pdf}
    \caption{}
    \label{fig:competition_distributions:weakest}
\end{subfigure}
\begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio=true]{competition_distribution_medium_bad.pdf}
    \caption{}
    \label{fig:competition_distributions:weak}
\end{subfigure}
\\
\begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio=true]{competition_distribution_medium_good.pdf}
    \caption{}
    \label{fig:competition_distributions:medium}
\end{subfigure}
\begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio=true]{competition_distribution_best.pdf}
    \caption{}
    \label{fig:competition_distributions:strong}
\end{subfigure}
\caption{Different inhibitory ($YY$) weights distributions: weakest competition (\subrefbf{fig:competition_distributions:weakest}), weak competition (\subrefbf{fig:competition_distributions:weak}), medium competition (\subrefbf{fig:competition_distributions:medium}), and strong competition (\subrefbf{fig:competition_distributions:strong})}
\label{fig:competition_distributions}
\end{figure}

@@

During the optimization of the models, the accuracy of a large number of LCSNNs was measured with various hyperparameter configurations (Fig. \refbf{fig:hyperparams}). It can be seen that configurations with high accuracy (dark lines) are localized in a manner that encourages negative weight accumulation for larger time differencies $t_{post} - t_{pre}$ between post- and pre- spikes (for the Anti-STDP weight update rule).

@@

Competition training makes it possible to achieve an accuracy slightly (by 2 \%) higher than the accuracy of a network of the same configuration, but with a fixed competition with inhibitory weights equal to $ -50 $ (Table \refbf{results}, №4 and №5).

In addition, inhibitory weights clamping (during training) experiments were carried out. It turned out that the inhibitory weights in the entire range of their variation are important for the effective LCSNN functioning, since both the upper and the lower clamps negatively affect the classification accuracy (Fig. \refbf{fig:compe_clamp}). This is due to the fact that high competition contributes to greater specialization of neurons and therefore is useful, and low competition allows neurons to cooperate and recognize digits together (including ensuring the accumulation of more statistics for calibrating the voites of neurons). Thus, the cooperation of neurons in the form of low competition between them is not less important than strong competition between neurons specialized on rather different image classes. This develops the idea of \parencite{MaxActiv1} work, in which the similar conclusion was made for inhibitory connections trained by terms of BCM-like local rules.

\begin{figure}[H] 
\centering 
\begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio=true]{competition_distribution_clamp_low.pdf}
    \caption{}
    \label{fig:compe_clamp:low_distr}
\end{subfigure}
\begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio=true]{competition_distribution_clamp_high.pdf}
    \caption{}
    \label{fig:compe_clamp:high_distr}
\end{subfigure}
\\
\begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio=true]{weights_XY_clamp_low.pdf}
    \caption{}
    \label{fig:compe_clamp:low_weights}
\end{subfigure} 
\begin{subfigure}{0.3\textwidth} 
    \includegraphics[width=\textwidth,keepaspectratio=true]{weights_XY_clamp_high.pdf}
    \caption{}
    \label{fig:compe_clamp:high_weights}
\end{subfigure}
\caption{Impact of limiting the values of inhibitory weights on accuracy. \subrefbf{fig:compe_clamp:low_weights}, \subrefbf{fig:compe_clamp:high_weights} show $XY$ weights and Figures \subrefbf{fig:compe_clamp:low_distr}, \subrefbf{fig:compe_clamp:high_distr} show the inhibitory ($YY$) weights distributions. The inhibitory weights were clamped (during training) to $(-\infty, -100]$ \subrefbf{fig:compe_clamp:low_distr},\subrefbf{fig:compe_clamp:low_weights} and to $[-350, 0]$ (\subrefbf{fig:compe_clamp:high_distr},\subrefbf{fig:compe_clamp:high_weights}). The accuracy of the network with the same hyperparameters, but without the competition limitation is 82.3\%, which is significantly higher than the accuracies of the networks presented in this figure. Its weights $ XY $ are shown in Fig. \refbf{fig:best_competition_XY}, and the distribution of its inhibitory weights is shown in Fig. \refbf{fig:competition_distributions:strong}. In both cases the networks develop more blurred features comparing to the network with unrestricted competition in Fig. \refbf{fig:competition_distributions:strong}
} 
\label{fig:compe_clamp}
\end{figure}

\section{Discussion} 
This study demonstrates that a locally connected network is a promising spiking neural network architecture suitable for efficient implementation on a specialized neuromorphic hardware \parencite{Emelyanov_2019, SURAZHEVSKY2021110890}. Indeed, the ability to quickly plateau the learning curve allows the LCSNN to be used for training on relatively small training samples (3000-5000). The use of an additional activity interpretation algorithm, such as a linear classifier, can improve the efficiency of the network. Note that in this case the very features of the objects being recognized are learned by the network in an unsupervised fashion. The question about the possibility of obtaining similar results without using supervised learning to interpret network activity, remains open.

The main advantage of LCSNN, as it has been shown, is that the locally connected network outperforms the convolutional in accuracy with approximately the same or slightly larger number of parameters. Presumably, this is due to the richer statistics of patches (receptive fields) that neurons within each channel respond to, compared to one receptive field per channel in the convolutional architecture. The price for this is the increased number of network parameters, but it's not a critical factor due to the local (as opposed to fully connected) topology of interneural contacts. The local architecture also allows to save information about the localization of the selected feature in space and therefore is a reasonable choice for the implementation of unsupervised trained SNNs. This choice also determines its overall efficiency (in terms of performance, accuracy and power consumption at a certain density of elements) in a hardware implementation, in which each weight is usually represented by one or more physical elements (mainly SRAM cells \parencite{TrueNorth, Loihi} or memristors \parencite{Li_2018, DEMIN202164}).
The training of inhibitory connections slightly increases the accuracy of the LCSNN-based algorithm, but it's not so crusial as to justify the use of such an expensive operation as the introduction and training of a significant number of inhibitory connections with variable values. At the same time, this conclusion is probably only valid only for small networks, like the one studied in this paper. In our opinion, the selection of optimal algorithms and parameters for learnable competition deep SNNs deserves a more thorough study, in which the balance of competition and cooperation of neurons within each layer can lead to the formation of clusters composed of supplied visually similar images. This contributes to building a semantical hierarchy of learned features and their combinations, up to the higher-level images \parencite{10.1007/978-3-030-30425-6_30}. Potentially, such a hierarchy of images can lead to a significant increase in the quality metrics.
